---
title: "Gov1005 Final Class Project"
author: "Hudson Miller"
date: "4/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

``` {r, include=FALSE, echo=FALSE}
library(pdftables)
library(readr)
library(tidyverse)
library(janitor)
library(readxl)
library(ggplot2)
library(dplyr)
library(plotly)
library(ggthemes)
```

```{r, CH1: Cleaning Active Duty Data on Race, echo=FALSE, include=FALSE}

# I pulled the following information from the Office of the Under Secretary of
# personal readiness. They issues an annual report on manpower and readiness for
# the armed forces. The most recent release is for FY2017. I used the PDFTables
# packages and https://pdftables.com/blog/convert-pdf-to-excel-r to pull tables
# from the report and convert them to excel files. Orginal source of the
# material was at https://www.cna.org/pop-rep/2016/summary/summary.pdf I am
# working on deciding what I want to combine based on what presentation medium I
# choose for each element of the project. I have a lot of data for education and
# military demographics but I need to decided what to do with it.

# This dataset has stats on the breakdown by race of people entering the military and has total numbers. loks at 2003-2017.

FY2017_DOD_Acc <- read_excel("Mil_data/FY2017 DOD.xlsx", skip = 3, col_types = "numeric") %>%
  clean_names() %>%
  slice(1:15) %>% rename(Year = year,
         Total_Active_Duty_White = white_2,
         Total_Active_Duty_Black = black_3,
         Total_Active_Duty_AIAN = aian_4,
         Total_Active_Duty_Asian = asian_5,
         Total_Active_Duty_NHPI = nhpi_6,
         Total_Active_Duty_Two_or_More = two_or_more,
         Total_Active_Duty_Unkown = unknown,
         Total_Active_Duty = total_9,
         Total_Civilian_White = white_10,
         Total_Civilian_Black = black_11,
         Total_Civilian_AIAN = aian_12,
         Total_Civilian_Asian = asian_13,
         Total_Civilian_NHPI = nhpi_14,
         Total_Civilian_Two_or_More = two_or_more_2,
         Total_Civilian = total_16)
```

```{r, CH2: Historical DOD Race Data, echo=FALSE, include=FALSE}

# This data covers accessions or new additions who are non prior service recruits. It 

DOD_Historical_Racial_Data_ <- read_excel("Mil_data/DOD Historical Racial Data %.xlsx") %>% clean_names()

```

```{r, CH3: DOD Race Data by Branch, echo=FALSE, include=FALSE}

# These two datasets are similar to the previous one but break it down by branch
# to give a more detailed picture. The first is the number of people per group
# and the second is the percent of the total for each group.

DOD_Acc_Branch_Numbers <- read_excel("Mil_data/DOD Acc Numbers (2003-17).xlsx.xlsx") %>% clean_names()

DOD_Acc_Branch_Percent_unkrm <- read_excel("Mil_data/DOD Acc Race Percent_cln.xlsx") %>% clean_names()

# These two datasets are similar to the previous two but show the current racial
# demographics of enlisted service members.

DOD_Enl_Branch_Numbers <- read_excel("Mil_data/DOD Enlisted Numbers (2003-17).xlsx") %>% clean_names()

DOD_Enl_Branch_Percent <- read_excel("Mil_data/DOD Enlisted Race Percent.xlsx") %>% clean_names()

# This dataset breaks down accession by region of the US.

DOD_race_by_region <- read_excel("Mil_data/DOD race by region.xlsx") %>% clean_names()

```

```{r, CH4: Cleaning Active Duty Data on Race with Percents, echo=FALSE, include=FALSE}

# I seperated the original data set into 2 files, one with number and the other
# with percentages to make it easier to use. I was able to pull the file from
# https://prhome.defense.gov/Portals/52/Documents/MRA_Docs/MPP/AP/poprep/2017/Appendix%20D%20-%20(Historical%20Component).pdf
# and it shows the stats on racial representation in the miltiary from 2003 to
# 2017.

FY2017_DOD_Acc_Percent <- read_excel("Mil_data/FY2017 DOD (%).xlsx") %>%
  clean_names() %>% 
  rename(Year = year,
         Acitve_Duty_White = x2,
         Acitve_Duty_Black = black_3,
         Acitve_Duty_AIAN = aian_4,
         Acitve_Duty_Asian = asian_5,
         Acitve_Duty_NHPI = nhpi_6,
         Acitve_Duty_Two_or_More = two_or_more,
         Acitve_Duty_UNKWN = unknown,
         Civilian_White = white,
         Civilian_Black = black_10,
         Civilian_AIAN = aian_11,
         Civilian_Asian = asian_12,
         Civilian_NHPI = nhpi_13,
         Civilian_Two_or_More = two_or_more_2)

```

```{r, CH5: Cleaning College Data on Race, echo = FALSE, include = FALSE}

# Racial and Gender numbers and percentages for college enrollment from selected
# dates between 1976 and 2004.All of the following files were taken from the
# National Center for Education Statistics at
# https://nces.ed.gov/programs/digest/d18/tables/dt18_306.10.asp. These 3 files
# provide overall enrollment statistics for US colleges and universities. They
# breakdown applicants, admitted students, and enrolled students by gender and
# by states. They also provide stats on racial statistics for undergrad and
# graduate universities. I have cleaned the names and changed certain names to
# make the data easier to use. 

# This data looks at the racial demographics on selected years from 1976 to 2017 and gathered from NCES.

Fall_enr_76_18_pct <- read_excel("Education_data/Fall_enr_76-18_pct.xlsx") %>% clean_names() %>% mutate(other=(asian_pi+aina+tom)) %>% select(1:4, 8)

Fall_enr_76_18_numbers <- read_excel("Education_data/Fall_enr_76-18_numbers.xlsx") %>% clean_names() %>% mutate(other=(asian_pi+aina+tom)) 

```

```{r, CH6: Cleaning College Data on Gender by State, echo = FALSE, include = FALSE}

ED_Gender_States_2015 <- read_excel("Education_data/ED_Gender_States_2015.xlsx") %>%
  rename(Area = 'State or jurisdiction',
  Total_Applicants = `Total1 _APP`,
  Male_Applicants = Men_APP,
  Female_Applicants = Women_APP,
  Total_Admitted = Total_Adm,
  Male_Admitted = Men_Adm,
  Female_Admitted = Women_Adm,
  Total_Enrollment = Total_Enr,
  Male_Enrollment = Men_Enr,
  Female_Enrollment = Women_Enr,
  Total_Enrollment_Full_Time = Total_Enr_FT,
  Male_Enrollment_Full_Time = Men_Enr_FT,
  Female_Enrollment_Full_Time = Women_Enr_FT,
  Total_Enrollment_Part_Time = Total_Enr_PT,
  Male_Enrollment_Part_Time = Men_Enr_PT,
  Female_Enrollment_Part_Time = Women_Enr_PT)

```

```{r, CH7: Cleaning College Data on Gender, echo = FALSE, include = FALSE}

ED_Gender_2018 <- read_excel("Education_data/ED_Gender_2018.xlsx") %>%
  clean_names 

```

```{r, CH8: Tidying Data Figure 1, include = FALSE, echo = FALSE}

# First I want to make sure my data is easy to work with for my graphic. I am
# taking the DOD data on percentages of each race over a period of time and
# transposing it so that the years match up with the education data. I added a
# group column that shows which group it is from so that I can tidy the data.

Select_modern_DOD <-DOD_Acc_Branch_Percent_unkrm %>%
  select(1:25)

# Once the data matched I combined the two datasets to have a greater amount of
# time covered. 

combined_dod_pct <- rbind(Select_modern_DOD, DOD_Historical_Racial_Data_)

# This was not the best way to do this but I could not find a faster way to so I
# had to. I separated by each group and then created a new column that allowed
# me to differentiated each group easily whenever I am working with the data.
# This made it a lot easier moving forward. Once all the groups were added, I
# combined that data again and had to make a few small changes so that it
# matched up with the data for undergraduate students.

Navy_PCT <- combined_dod_pct %>%
  select("year"=fiscal_year,
         "white"=navy_white,
         "black"=navy_black,
         "hispanic"=navy_hispanic,
         "other"=navy_non_hispanic_other) %>%
  mutate(group = "navy")

Army_PCT <- combined_dod_pct %>%
  select("year"=fiscal_year,
         "white"=army_white,
         "black"=army_black,
         "hispanic"=army_hispanic,
         "other"=army_non_hispanic_other) %>%
  mutate(group = "army")

USMC_PCT <- combined_dod_pct %>%
  select("year"=fiscal_year,
         "white"=usmc_white,
         "black"=usmc_black,
         "hispanic"=usmc_hispanic,
         "other"=usmc_non_hispanic_other) %>%
  mutate(group = "USMC")

AF_PCT <- combined_dod_pct %>%
  select("year"=fiscal_year,
         "white"=air_force_white,
         "black"=air_force_black,
         "hispanic"=air_force_hispanic,
         "other"=air_force_non_hispanic_other) %>%
  mutate(group = "Air Force")

DOD_PCT <- combined_dod_pct %>%
  select("year"=fiscal_year,
         "white"=total_dod_white,
         "black"=total_dod_black,
         "hispanic"=total_dod_hispanic,
         "other"=total_dod_non_hispanic_other) %>%
  mutate(group = "DOD")

Civ_PCT <- combined_dod_pct %>%
  select("year"=fiscal_year,
         "white"=total_civilian_18_24_white,
         "black"=total_civilian_18_24_black,
         "hispanic"=total_civilian_18_24_hispanic,
         "other"=total_civilian_18_24_non_hispanic_other) %>%
  mutate(group = "civilian")

Fig1_DOD_data <- rbind(Navy_PCT, Army_PCT, USMC_PCT, AF_PCT, DOD_PCT, Civ_PCT)

Fig1_Enr_data <- Fall_enr_76_18_pct %>%
  mutate(group = "college")

Fig1_joined_data <- rbind(Fig1_Enr_data, Fig1_DOD_data)

Fig1_joined_data$white <- Fig1_joined_data$white %>%
  as.numeric()

Fig1_joined_data$black <- Fig1_joined_data$black %>%
  as.numeric()

Fig1_joined_data$hispanic <- Fig1_joined_data$hispanic %>%
  as.numeric()

Fig1_joined_data$other <- Fig1_joined_data$other %>%
  as.numeric()

```

```{r, CH9: Figure 1 Data, echo=FALSE}

# This is the chunk where I saved the data that I need for my first ineractive
# figure.

Fig1_joined_data_bg <- Fig1_joined_data %>%
  filter(group == "DOD" | group == "college" | group == "civilian") %>%
  na.omit()

save(Fig1_joined_data_bg, file = "FPShiny/Fig1_joined_data_bg.Rdata")

Fig_2 <- Fig1_joined_data %>%
  pivot_longer(names_to = "race",
               values_to = "amount",
               cols = c(-year, -group)) %>% na.omit()
               
```


```{r, CH10: Figure 2 (Gender), echo=FALSE}

Fall_1976_2017_num <- read_excel("Education_data/Fall_1976_2017_num.xls")

Fall_1976_2017_pct <- read_excel("Education_data/Fall_1976_2017_pct.xlsx")

Gen_Race_pct_enr <- Fall_1976_2017_pct %>%
  pivot_longer(names_to = "Year",
               values_to = "Percent",
               cols = c(-Race, -Gender)) %>% filter(Race == "White" | Race == "Black" | Race == "Hispanic")

Gen_Race_pct_enr$Percent <- Gen_Race_pct_enr$Percent %>% round(1)

Gen_Race_pct_Male <- Gen_Race_pct_enr %>% filter(Gender == "Male") 

Gen_Race_pct_Female <- Gen_Race_pct_enr %>% filter(Gender == "Female") 

save("Gen_Race_pct_Male", file="FPShiny/Gen_Race_pct_Male.Rdata")

Gen_Race_pct_Female %>%  ggplot(aes(Year, Percent, color=Gender, fill=Race)) +
    geom_bar(stat = "identity", position = "dodge") + theme_bw() + labs(x="Year")

save("Gen_Race_pct_Female", file="FPShiny/Gen_Race_pct_Female.Rdata")
```

# ggplot(G1_Data, aes(x = Fiscal_Year, y = White, color = Group)) + geom_line() + theme_grey() + labs(title = "Percentage of Population Who are White", subtitle = "Comparing Civilians(ages 18-24 years old), Active Enlisted Department of Defense Service Members, and Enrolled Undergraduate Students", caption = "Sources: National Center for Education Statistics and Department of Defense", color = "Population group")
# 
# ggsave("Graphic_1.png", path = "FP_SHINY")

```{r, CH11: Graphic, echo = FALSE}

# DOD_Enl_by_Branch_Percent
# Civ_Race_Pct_G1
```


## Could the military learn better gender representation from colleges and could colleges learn better racial diversity from the military?

### About the Project:
#### The purpose of this project is to compare and analyze the difference in racial and gender diversity between the United States Armed Forces and US colleges and universities. The military has become more racial diverse in last few decades but despite this trend towards diversity, there has not been a surge of females to the armed forces. Likewise, US colleges and universities have seen higher numbers of women enroll but they still lack racial diversity. 

### Methodology:
#### This project will use college enrollment data gathered from National Center for Education Statistics and military demographic statistics gathered by the Office of the Under Secretary of Defense for Personnel and Readiness and published in the Fiscal Year 2017 Population Representation in the Military Service. In order to compare the new fields, I will first display the racial and gender breakdowns for each field. I will then see if those racial or gender statistics can be explained by where the majority of applicants for the military or college are located.

## https://hudson-miller.shinyapps.io/FP_SHINY/
# DOD Data


#CH 1
View(FY2017_DOD_Acc)
# CH 2
View(DOD_Historical_Racial_Data_)
# CH 3
View(DOD_Acc_Branch_Numbers)
View(DOD_Acc_Branch_Percent_unkrm)
View(DOD_Enl_Branch_Numbers)
View(DOD_Enl_Branch_Percent)
View(DOD_race_by_region)
# CH 4
View(FY2017_DOD_Acc_Percent)

# College Data

# CH 5
View(Fall_enr_76_18_pct)
View(Fall_enr_76_18_numbers)

# CH 6
View(ED_Gender_States_2015)

# Ch 7
View(ED_Gender_2018)

# Ch 8


